---
title: "chapter5"
output: html_document
---
#Chapter 5
###data
```{r}
human<-read.csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep=',', header= TRUE)

```

###overview and summary of the data
```{r}
pairs(human)
summary(human)
```



It seems that Edu.Exp is positvely correlated to Life.Exp and negatively to Mat.Mor.
Mat.Mor seems negatively correlated to Life.Exp. 

###Try to find out the most essential variabls through dimentionality reduction with PCA. 

###the unstandardized variables

```{r}
pca_human_unstd<-prcomp(human)
us<-summary(human)
biplot(pca_human_unstd, cex=c(0.8,1), col=c("grey40","deeppink2"), xlab="PC1", ylab="PC2")
```

###the standized variables
```{r}
human_std<-scale(human)
```
###print out summaries of human_std
```{r}
summary(human_std)
```
###perform principal component analysis(with the SVD method)
```{r}
pca_human<-prcomp(human_std)
pca_human
```
###draw a biplot of the principan component representation and the original variables
```{r}
s<-summary(pca_human)
```
###rounded percentages of variance captured by each PC
```{r}
pca_pr<-round(1*s$importance[2,], digits=1)
```
###print out the percentages of variance
```{r}
pca_pr
```
###create object pc_lab to be used as axis labels
```{r}
pc_lab<-paste0(names(pca_pr), "(", pca_pr, "%)")
```
###draw a biplot
```{r}
biplot(pca_human, cex=c(0.8,1), col=c("grey40", "deeppink2"), xlab=pc_lab[1], ylab=pc_lab[2])
```

Unstandaized variables looks out of ranges because PCA is sensitive to the relatives scallings of the orginal variables and assumes that feature with larger varianace are more important than variables with smaller variabnce. 


###Multiple Correspondence Analysis with Tea dataset

```{r}
library(FactoMineR)
library(ggplot2)
library(dplyr)
library(tidyr)
data("tea")
```

###keep column names in the dataset
```{r}
keep_columns<-c("Tea", "How","how","sugar","where","lunch")
```
###select the 'keep_columns' to create a new dataset
```{r}
tea_time<-select(tea, one_of(keep_columns))
```
###summaries and structure of the data
```{r}
summary(tea_time)
str(tea_time)
```
###visualize the dataset
```{r}
gather(tea_time)%>%ggplot(aes(value))+facet_wrap("key",scales="free")+geom_bar()
```

###Multiple Corresondence Analysis
```{r}
mca<-MCA(tea_time, graph=FALSE)
```
###summary of the model
```{r}
summary(mca)
```
###visualize MCA
```{r}
plot(mca, invisible=c("ind"), habillage="quali")
```
###biplot
biplot()